---
title: "Data Project"
description: |
  Abalone Data from Kaggle 
author:
  - name: Tracy Morgan
    

date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Data Set: Abalone from Kaggle 
ABALONE:  a marine snail.  You can read about it here in Wikipedia!  
Data Description: 4.177 records of measurements, age, and sex of these oyster-like creatures called Abalone.  
Descriptive Stats: There are some large specimen outliers included in the data set.  
Machine Learning Model:
Run Model and Find Results:
Describe Results:  


```{r}
library(caret)
library(ggplot2)
library(tidyverse)
library(readxl)
library(psych)
library(knitr)

```
```{r}
library(readr)
abalone <- read_csv("abalone.csv")
View(abalone)

```
```{r,echo=TRUE}
list(abalone)
```

```{r}
   # Create a column name from abalone.txt to get a description of each record value
    colnames(abalone) <- c('Sex', 'Length', 'Diameter', 'Height', 'WholeWeight', 'ShuckedWeight',
                           'VisceraWeight', 'ShellWeight', 'Rings')

```
# Get a summary of data 
```{r}
 summary(abalone)
```
# Get type of data
```{r}
  sapply(abalone, class)
```

# Plotting Age Linear Models
According to correlation results more correlation is between ShellWeight and Rings and Length and Rings

```{r}
qplot(Length, Rings, data=abalone, color=Sex) + geom_point() + geom_smooth(method="lm")
```
  
```{r}
qplot(ShellWeight, Rings, data=abalone, color=Sex) + geom_point() + geom_smooth(method="lm")
```
# Comparison linear models
We will compare some models in order to select the best
   
```{r}
    modelFitAll <- lm(Rings ~ ., data = abalone) 
    modelFitMultiple <- lm(Rings ~ ShellWeight + Length, data = abalone)   
    modelFitShellWeight <- lm(Rings ~ ShellWeight, data = abalone) 
    plot(resid(modelFitShellWeight), col="Blue")
```
```{r}
anova(modelFitMultiple, modelFitShellWeight, test="Chisq")
```
```{r}
    anova(modelFitAll, modelFitShellWeight, test="Chisq")
```

# Seems than better model is taking into account all variables to predict Rings
```{r}
summary(modelFitAll)
```
```{r}
   plot(modelFitAll)
```
```{r}
   # Rings linear model
    inTrainRings <- createDataPartition(y=abalone$Rings, p=0.75, list=FALSE)
    trainingRings <- abalone[inTrainRings,]
    testingRings <- abalone[-inTrainRings,]
    set.seed(1234)
    modelFitRings <- train(Rings~., data=trainingRings, method="lm")
    #predict(modelFitRings, data = testingRings)
    sqrt(sum((modelFitRings$fitted - trainingRings$Rings)^2))
```
```{r}
sqrt(sum((predict(modelFitRings, newdata = testingRings) - testingRings$Rings)^2))
```
    


