---
title: "Data Project "
description: |
  Predicting House Prices using R 
author:
  - name: I found this regression anaysis on Kaggle, written by Pradeep Tripathi 
    url: "https://www.kaggle.com/pradeeptripathi/predicting-house-prices-using-r"

date: "`r Sys.Date()`"
output: distill::distill_article



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Data Set: 

My Goal: 


Really!  I'm lost so I'm switching to an easier data set.  
Abalone is what another student is working on so I'm going to try to look at this data set: 

ABALONE:  a marine snail.  You can read about it here in Wikipedia!  

Great, I'm looking at house prices and then  I am changing to snails. 
```{r}
library(caret)
library(ggplot2)

```
```{r}
library(readr)
abalone <- read_csv("abalone.csv")
View(abalone)
```


```{r}
   # Create a column name from abalone.txt to get a description of each record value
    colnames(abalone) <- c('Sex', 'Length', 'Diameter', 'Height', 'WholeWeight', 'ShuckedWeight',
                           'VisceraWeight', 'ShellWeight', 'Rings')
```

```{r}
   # Get a summary of data 
    summary(abalone)
```
```{r}
    # Get type of data
    sapply(abalone, class)
```
```{r}
 qplot(Length, Rings, data=abalone, color=Sex) + geom_point() + geom_smooth(method="lm")
```



